Towards a Neural Co-Processor Which Restores Dexterity After Stroke: Modeling a Proof-of-Concept

### 2021-12-12 paper/article.tex
Paper progress

\subsubsection{EN training}
For the EN to be useful, it must accurately predict the effect of stimulation
produced by the CPN. In addition to that, backpropagating through it must yield
gradients which train the CPN to output more useful stimulation. We discovered that
this latter property does not necessarily occur simply by virtue of the former.
We have found that an EN can be trained to extreme levels of predictive power, even
on random stimulation, and to orders of magnitude lower loss than the task loss,
while at the same time backpropagation through it yields gradients which are not
useful for training the CPN. As a result, if the EN is not trained in specific ways,
the CPN exhibits unstable training. We have thus far been unable to discover a way
to measure and optimize for this second property directly, though pursuing it may be a
fruitful subject of future applied ML research. Instead of attempting to train an EN
directly to exhibit this property, we discovered a training regime that indirectly,
but reliably, makes it occur.

@@

\subsubsection{Improving learning efficiency with data reuse}
TODO. May drop this section if I can't get learning stabilized with this approach.

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{results_standin.png}
\caption{Loss and class separation vs training epoch}
\centering
\label{fig:results}
\end{figure}

@@

TODO: summary of conclusions that will go here:
\begin{itemize}
	\item Task performance improves drastically. In all cases, performance is still improving slowly when I stop the simulation.
		\begin{itemize}
			\item 90\% on connection lesion, both with and without co-adaptation
			\item M1 50\%: 81\% recovery with and without co-adpation. The co-processor is having trouble finding a better solution than this, due to loss in brain functionality. Increasing the rate of brain co-adaptation may show a better solution.
			\item AIP 50\%: 53\% recovery w/co-adaptation so far; still running. The loss was very low to start with, and recovery depends heavily on co-adaptation, so training is slow. Without co-adaptation, the co-processor cannot learn well, because too much encoding machinery is lost.
			\item Waiting on pre-recovery result. Unclear if the co-processor will help or not.
		\end{itemize}
	\item Initially, performance improves orders of magnitude more quickly than later. 50-75\% recovery during this phase. This appears to be the co-processor learning the task structure (e.g. not to move the hand until after the hold signal is lifted).
	\item Fine tuning takes a far longer time
	\item This may have some connection to object class separation; show class sep metric side-by-side. (I have the data in hand, but haven't tried to correlate it to recovery yet).
	\item Training efficiency analysis. The majority of epochs (80\%; depends on experiment type) are EN training epochs, so most efficiency gains will be there.
	\item Can co-adapt with the brain, as it changes
	\item Compare co-adaptive case vs not. Some tasks will be impossible to accomplish without the two cooperating, if enough cortex has been lost. (waiting for new results to confirm)
	\item Show pre-recovery results; co-processor will hopefully improve beyond what recovery can achieve. Initial result: the co-adaptive case found a better solution. The co-processor couldn't improve yet on the pre-recovery case. The solution it found was (basically) to ignore the inputs.
\end{itemize}

@@

Our co-processor design adapts well to a variety of simulated lesion types, reducing task loss
50-90\% across our various experiments. To work this well, the co-processor needed to adapt to
the long-running dynamics of the stimulated network, as well as the long-running effects of
stimulation. In some experiments we required it to additionally adapt to non-stationarity in
the neural circuit, which was actively changing at the same time the co-processor was learning.
It achieved that as well, co-adapting with the simulated brain to improve external task
performance.  TODO: pre-recovery result.

>>>>

\subsubsection{EN training}
For the EN to be useful, it must accurately predict the effect of stimulation
produced by the CPN. In addition to that, backpropagating through it must yield
gradients which train the CPN to output more useful stimulation. We discovered that
this latter property does not occur simply by virtue of the former.
We have found that an EN can be trained to high levels of predictive power, even
on random stimulation, and to orders of magnitude lower loss than the task loss,
while at the same time backpropagation through it yields gradients which are not
useful for training the CPN. As a result, if the EN is not trained in specific ways,
the CPN exhibits unstable training. We have thus far been unable to discover a way
to measure and optimize for this second property directly, though pursuing it may be a
fruitful subject of future applied ML research. Instead of attempting to train an EN
directly to exhibit this property, we discovered a training regime that indirectly,
but reliably, makes it occur.

@@

@@

\begin{figure}[h]
\includegraphics[width=\textwidth]{training_results.png}
\caption{Loss and class separation vs training epoch}
\centering
\label{fig:results}
\end{figure}

@@

TODO: summary of conclusions that will go here:

TODO:
* AIP + coadaptation shows that we can adapt to changes in the input mappings.
* M1 + coadaptation shows we can adapt to changes in the output mappings
* Connection lesion shows we can act as a bridge

\begin{itemize}
	\item Task performance improves drastically. In all cases, performance is still improving slowly when I stop the simulation.
		\begin{itemize}
			\item 90\% on connection lesion, both with and without co-adaptation
			\item M1 50\%: 81\% recovery with and without co-adpation. The co-processor is having trouble finding a better solution than this, due to loss in brain functionality. Increasing the rate of brain co-adaptation may show a better solution.
			\item AIP 50\%: 53\% recovery w/co-adaptation so far; still running. The loss was very low to start with, and recovery depends heavily on co-adaptation, so training is slow. Without co-adaptation, the co-processor cannot learn well, because too much encoding machinery is lost.
			\item Waiting on pre-recovery result. Unclear if the co-processor will help or not.
		\end{itemize}
	\item Initially, performance improves orders of magnitude more quickly than later. 50-75\% recovery during this phase. This appears to be the co-processor learning the task structure (e.g. not to move the hand until after the hold signal is lifted).
	\item Fine tuning takes a far longer time
	\item This may have some connection to object class separation; show class sep metric side-by-side. (I have the data in hand, but haven't tried to correlate it to recovery yet).
	\item Training efficiency analysis. The majority of epochs (80\%; depends on experiment type) are EN training epochs, so most efficiency gains will be there.
	\item Can co-adapt with the brain, as it changes
	\item Compare co-adaptive case vs not. Some tasks will be impossible to accomplish without the two cooperating, if enough cortex has been lost. (waiting for new results to confirm)
	\item Show pre-recovery results; co-processor will hopefully improve beyond what recovery can achieve. Initial result: the co-adaptive case found a better solution. The co-processor couldn't improve yet on the pre-recovery case. The solution it found was (basically) to ignore the inputs.
\end{itemize}

@@

Our co-processor design adapts well to a variety of simulated lesion types, reducing task loss
50-90\% across our various experiments. To work this well, the co-processor needed to adapt to
the long-running dynamics of the stimulated network, as well as the long-running effects of
stimulation. In some experiments we required it to additionally adapt to non-stationarity in
the neural circuit, which was actively changing at the same time the co-processor was learning.
It achieved that as well, co-adapting with the simulated brain to improve external task
performance. Finally, it adapted to a simulated brain which had already undergone some amount of
lesion recovery. In that case, the co-processor successfully identified the information upstream
from the lesion which was necessary to stimulate the motor cortex downstream.
